In-Band Synchronization for\\Distributed SDN Control Planes

### 2015-09-01 reviews/response-rev1-stefan.txt

Liron, Petr, Stefan

Reviewer 1
==========

> Overall, the work seems useful and correct.
> My main concern is scope.
> The paper is essentially a note on OpenFlow hacking, and thus might be
> of limited interest to many CCR readers. However, OpenFlow and similar
> SDN approaches have seen sufficient adoption and research interest in
> recent years that there is utility in this paper appearing in a venue with
> broad reach.

Thank you very much for your feedback. Indeed, we hope and believe that
our work is of relevance beyond specific Openflow versions. In our revision,
we aim to put the fundamental aspects and the benefit of inband synchronization
into a broader context. Having that said, in our revision we also report
in more detail on our proof-of-concept implementation.

@@

> pg 3 col 2: "a bundle allows to conditionally"

Fixed.

> In Example 1, why not have the two controllers communicate
> Example 2 says bundling and checkoverlap don't help modify. But, Section 2
> says that the semantics of addflow is to modify if a flow with exactly
> the same priority and match exist

In our revision, we restructured these parts and hope we found
a more intuitive way to put this into perspective.

### 2015-09-01 ccr-rev1/inband-sync.tex
dsaf

\begin{abstract}
Control planes of forthcoming Software-Defined Networks (SDNs)
will be \emph{distributed}: to ensure availability and fault-tolerance,
to improve load-balancing, and to reduce overheads,
modules of the control plane %operating on the data plane 
should be physically distributed.
However, %a distributed control plane also introduces
%new challenges,
in order to guarantee \emph{consistency} of network operation, actions
performed on the data plane by different controllers may need to 
be \emph{synchronized},  which is a
 nontrivial task.
In this paper, we propose a synchronization
framework for control planes  based on the well-known concept of
\emph{transactions}.
In particular, we present a powerful interface 
to implement transactions \emph{in-band}, on the 
switch itself
and using
the data plane configuration as a shared memory.
We argue that this in-band approach is attractive as 
it keeps the failure scope local and 
does not require additional 
out-of-band coordination mechanisms.
This allows us 
to realize fundamental consensus primitives in the presence
of controller failures, and we discuss applications
for 
 consistent policy composition and fault-tolerant control-planes.
Interestingly, 
by leveraging the match-action
paradigm to realize simple key-value stores, 
we can implement our synchronization framework in
today's standard Openflow protocol, 
and we report on
a proof-of-concept accordingly.
\end{abstract}

@@

We propose to implement our synchronization constructs \emph{in-band},
on the switch.
An in-band implementation allows us to efficiently alleviate the problems related to
coordinating controllers via a control plane
out-of-band network, in the presence of asynchrony and failures.
Indeed, the inherent costs  of such protocols are often considered too high, both in
terms of the necessary computability assumptions about the underlying
system~\cite{FLP85}, and the high communication
complexity~\cite{Lam06}.
In contrast, our \emph{in-band} solution allows the
controllers to solve fundamental agreement tasks in just one message
exchange with the data plane, tolerating asynchrony and failures of
any number of controllers.

@@

This paper proposes mechanisms to solve such more general 
synchronization problems.

@@

First, %if employed without any additional features,
except for some limited constraint types (such as
overlapping flow spaces or insufficient free space),
a bundle does not provide any means to modify the switch configuration
based on application-specific conditions, which depend on the
current configuration,
However, as we have for example seen in our load-balancing
example, a support for more generic conditions under which configuration
updates
should or should not take place, is essential.
Moreover, it is sometimes desirable to react to conflicts in smarter ways than
by simply aborting an operation, e.g., by supporting \emph{conditional modifications}.

@@

In order to check if an identifier is claimed, a controller tries to add
a new entry setting the flag \texttt{\checko}, and using a different
match where the controller id is shifted $16$ bits to the left,
thereby inflicting a failure in case an entry with the same claimed value exists.
If the check succeeds, we delete the entry. In any case since the check entry contains the controller id, it doesn't conflicts with other check entries and there is no negative impact of delaying the deletion.
%Considering $m_1$ and $m_2$ from the example above,
%with respective actions $a_1=1$ and $a_2=2$,  they overlap
%with different actions iff $x_1=x_2$. Therefore, if $m_1$ is used in the \claimcheck after $m_2$ was used in the claim operation,
%a conflict is detected iff $x_1=x_2$, as desired.

@@

Let us now revisit our load-balancing application
from the beginning of the paper.
A simple strategy to balance load in our 2-link example
is given in Algorithm~\ref{alg:naive-lb}.
However, this algorithm fails in the presence of
multiple controllers, even if 
bundling and $\checko$ features are used:
the flows are defined over ``independent'' 
flow spaces.

@@

Let us consider a simple policy update protocol that
allows a set of  controllers to concurrently update switch policies, i.e., sets of
%[[PK policies have been defined in Background]]
effective flow entries, under the
condition that these rules can be \emph{composed} with the currently installed
policy.
Algorithm~\ref{alg:simple-update} uses the bundle feature to update
the switch configuration, where the configuration contains both the
policy and a meta-configuration memory address $\paddr$ that holds the \emph{policy identifier (pid)} .
In the algorithm, the controller first reads the currently installed
policy together with its {\pid},  applies the
$\textsc{\ufunc}()$ function to the current policy which results in a
set of update FlowMod commands and then tries to apply them
atomically together with a CAS operation on $\paddr$ replacing $\pid$
with $\pid+1$.
The latter ensures that if, concurrently, a new policy (with a different identifier) has been installed, the update
fails and takes no effect.
Algorithm~\ref{alg:simple-update} can therefore also be used
to render our na\"ive load-balancing Algorithm~\ref{alg:naive-lb}
consistent, without any modifications in the update function.

@@

We experimented with the low-level abstractions as well as the 
transactional policy update algorithm of our proof-of-concept, using
different workloads. 
The tests were conducted using a software switch, Open vSwitch (version 2.4.0) and virtual network simulator, Mininet.

For instance, we experimented with a setting where three processes
(controllers) issue thousands of concurrent policy updates
to the switch, where 
each update consists in replacing a single rule with another one.
When verifying the transactions, we could confirm that 
despite the concurrency, all transactions have been successfully completed.
Of course, due to the concurrency, multiple transactions also needed to abort,
due to conflicts.

>>>>

\begin{abstract}
Control planes of forthcoming Software-Defined Networks (SDNs)
will be \emph{distributed}: to ensure availability and fault-tolerance,
to improve load-balancing, and to reduce overheads,
modules of the control plane %operating on the data plane 
should be physically distributed.
However, %a distributed control plane also introduces
%new challenges,
in order to guarantee \emph{consistency} of network operation, actions
performed on the data plane by different controllers may need to 
be \emph{synchronized},  which is a
 nontrivial task.
In this paper, we propose a synchronization
framework for control planes  based on the well-known concept of
\emph{transactions}.
In particular, we present a powerful interface 
to implement transactions \emph{in-band}, on the 
switch itself
and using part of
the data plane configuration space as a shared memory.
We argue that this in-band approach is attractive as 
it keeps the failure scope local and 
does not require additional 
out-of-band coordination mechanisms.
This allows us 
to implement fundamental consensus primitives in the presence
of controller failures, and we discuss their applications
for 
 consistent policy composition and fault-tolerant control-planes.
Interestingly, 
by leveraging the match-action
paradigm to realize simple key-value stores, 
we can implement our synchronization framework in
today's standard Openflow protocol, 
and we report on
a proof-of-concept accordingly.
\end{abstract}

@@

We propose to implement our synchronization constructs \emph{in-band},
on the switch.
An in-band implementation allows us to efficiently alleviate the problems related to
coordinating controllers via a control plane
out-of-band network, in the presence of asynchrony and failures.
Indeed, the inherent costs  of such conventional \emph{out-of-band} 
protocols are often considered too high, both in
terms of the necessary computability assumptions about the underlying
system~\cite{FLP85}, and the high communication
complexity~\cite{Lam06}.
In contrast, our \emph{in-band} solution allows the
controllers to solve fundamental agreement tasks in just one message
exchange with the data plane, tolerating asynchrony and failures of
any number of controllers.

@@

This paper proposes mechanisms to solve these and more general 
synchronization problems.

@@

First, %if employed without any additional features,
except for some limited constraint types (such as
overlapping flow spaces or insufficient free space),
a bundle does not provide any means to modify the switch configuration
based on application-specific conditions, which depend on the
current configuration.
However, as we have seen in our load-balancing
example, a support for more generic conditions under which configuration
updates
should or should not take place, is essential.
Moreover, it is sometimes desirable to react to conflicts in smarter ways than
by simply aborting an operation, e.g., by supporting \emph{conditional modifications}.

@@

In order to check if an identifier is claimed, a controller tries to add
a new entry setting the flag \texttt{\checko}, and using a different
match where the controller id is shifted $16$ bits to the left,
thereby inflicting a failure in case an entry with the same claimed value exists.
If the check succeeds, we delete the entry. In any case since the check entry contains the controller id, it does not conflict with other check entries and there is no negative impact of delaying the deletion.
%Considering $m_1$ and $m_2$ from the example above,
%with respective actions $a_1=1$ and $a_2=2$,  they overlap
%with different actions iff $x_1=x_2$. Therefore, if $m_1$ is used in the \claimcheck after $m_2$ was used in the claim operation,
%a conflict is detected iff $x_1=x_2$, as desired.

@@

Let us now revisit our load-balancing application
from the beginning of the paper.
A simple strategy to balance load in our 2-link example
 fails in the presence of
multiple controllers, even if 
bundling and $\checko$ features are used:
the flows are defined over ``independent'' 
flow spaces.

@@

Let us consider a simple policy update protocol that
allows a set of  controllers to concurrently update switch policies, i.e., sets of
%[[PK policies have been defined in Background]]
effective flow entries, under the
condition that these rules can be \emph{composed} with the currently installed
policy.
Algorithm~\ref{alg:simple-update} uses the bundle feature to update
the switch configuration, where the configuration contains both the
policy and a meta-configuration memory address $\paddr$ that holds the \emph{policy identifier (pid)} .
In the algorithm, the controller first reads the currently installed
policy together with its {\pid},  applies the
$\textsc{\ufunc}()$ function to the current policy which results in a
set of update FlowMod commands and then tries to apply them
atomically together with a CAS operation on $\paddr$ replacing $\pid$
with $\pid+1$.
The latter ensures that if, concurrently, a new policy (with a different identifier) has been installed, the update
fails and takes no effect.
Algorithm~\ref{alg:simple-update} can therefore also be used
to render a na\"ive, single-controller load-balancing algorithm
consistent, without any modifications in the update function.

@@

We experimented with the low-level abstractions as well as the 
transactional policy update algorithm of our proof-of-concept, using
different workloads. 
The tests were conducted using a software switch, Open vSwitch (version 2.4.0) and
a virtual network simulator, Mininet.

For instance, we experimented with a setting where three processes
(controllers) issue thousands of concurrent policy updates
to the switch, where 
each update consists in replacing a single rule with another one.
When verifying the transactions, we could confirm that 
despite the concurrency, 100 \% of the transactions have been successfully completed.
Of course, due to the concurrency, multiple transactions also needed to abort,
due to conflicts.
